---
title: "Weighted expected binomial model"
author: "Anthony Aylward"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Prepare the data.

```{r}
library(chenimbalance)
total_reads <- rowSums(accb[, c("cA", "cC", "cG", "cT")])
data <- data.frame(
  total = total_reads,
  allelicRatio = sapply(
    1:nrow(accb),
    function(i) {
      accb[[paste("c", accb[["ref"]][[i]], sep = "")]][[i]] / total_reads[[i]]
    }
  )
)
head(data)
```

## Empirical distribution

Compute the empirical allelic ratio distribution

```{r}
binSize <- 40
bins <- pretty(0:1, binSize)
minN <- 6
maxN <- min(2500, max(data[["total"]]))
empirical <- empirical_allelic_ratio(
  data,
  bins,
  maxN = maxN,
  minN = minN,
  plot = TRUE
)
```

## Expected Binomial and Beta-Binomial distributions

Compute the weighted expected binomial distribution

```{r}
w <- weight_by_empirical_counts(data[["total"]])
d_combined_sorted_binned <- nulldistrib(
  w,
  minN = minN,
  binSize = binSize
)
```

Compute the sum of squared errors for the empirical distribution vs the 
weighted expected binomial distribution.

```{r}
sse <- sum((empirical - d_combined_sorted_binned[,2])^2)
sse
```

Choose the overdispersion parameter for the beta-binomial distribution

```{r}
w_grad <- graded_weights_for_sse_calculation(r_min = 0, r_max = 1, bins = bins)
overdispersion_details <- choose_overdispersion_parameter(
  w_grad,
  w,
  empirical,
  sse
)
head(overdispersion_details[["b_and_sse"]])
```

Generate a plot of the weighted expected binomial and weighted expected 
beta-binomial distributions overlaid on the empirical distribution

```{r}
plot_distributions(
  minN,
  maxN,
  bins,
  empirical,
  d_combined_sorted_binned,
  overdispersion_details[["e_combined_sorted_binned"]],
  yuplimit = 0.15
)
```

`overdispersion_details` is a list whose elements include the chosen value of 
`b` and the sum of squared errors.

```{r}
paste(
  "b_chosen =",
  overdispersion_details[["b_choice"]],
  ", SSE_chosen =",
  overdispersion_details[["sse"]]
)
```

Optimize the overdispersion parameter

```{r}
optimized_overdispersion_details <- optimize_overdispersion_parameter(
  w_grad,
  overdispersion_details[["b_and_sse"]],
  overdispersion_details[["b_choice"]],
  overdispersion_details[["sse"]],
  empirical,
  overdispersion_details[["counter"]],
  minN = minN,
  binSize = binSize
)
plot_distributions(
  minN,
  maxN,
  bins,
  empirical,
  d_combined_sorted_binned,
  optimized_overdispersion_details[["e_combined_sorted_binned"]],
  yuplimit = 0.15
)
```

Check the optimized value

```{r}
list(
  b = optimized_overdispersion_details[["b_choice"]],
  sse = optimized_overdispersion_details[["sse"]]
)
```

Plot the parameter search space

```{r}
b_and_sse <- b_and_sse[1:(optimized_overdispersion_details[["counter"]] + 2,]
plot(
  b_and_sse[order(b_and_sse[,1]),],
  type = "b",
  pch = 16,
  xlim = c(min(b_and_sse[,1]), max(b_and_sse[,1])),
  ylim = c(min(b_and_sse[,2]), max(b_and_sse[,2]))
)
par(new = TRUE)
plot(
  optimized_overdispersion_details[["b_choice"]],
  optimized_overdispersion_details[["sse"]],
  bty = "n",
  ylab = "",
  xlab = "",
  yaxt = "n",
  xaxt = "n",
  col = "red",
  pch = 8,
  xlim = c(min(b_and_sse[,1]), max(b_and_sse[,1])),
  ylim = c(min(b_and_sse[,2]), max(b_and_sse[,2]))
)
```
